---
title: "Election Day -Aaryan"
format: html
editor: Aaryan Bhatta
---

## Libraries

```{r}
library(tidyverse)
library(sf)
library(readxl)
library(RColorBrewer)
library(parallel)
```

# Data Sets

I utilized data sets, that helped me to visualize a map of political affiliations and the prediction turnout of registered voters. The turnout data set, I later used for calculating the turnout prediction for 2024's presidential election.

```{r}
map2024 <- st_zm(st_read("C:\\Users\\Aaryan D B\\Documents\\GitHub\\DSCI210-f24\\data\\maps\\precincts_2024.shp"))
voters.sos <-read_csv("voters_SOS.csv")
turnout <- read_excel("Turnout 2012-2023.xlsx")
```

# Map of Political Affiliations of Registered Voters

When creating this map, I combined the data of registered voters and the map data of Hamilton county. I managed to determine voters who were democrat, republican and unaffiliated. Based on that, I was able to calculate the proportion of the different affiliations. I this case, I wanted to look at the democrats proportion against the republican and I produced a map of it.

```{r}
voter.summary <- 
  voters.sos %>%
  mutate(PARTY_AFFILIATION = replace_na(PARTY_AFFILIATION, 'U')) %>% 
  group_by(PRECINCT_NAME) %>%
  summarize(total_voters = n(),
            un_voters = sum(PARTY_AFFILIATION == "U"), 
            rep_voters = sum(PARTY_AFFILIATION == "R"), 
            dem_voters = sum(PARTY_AFFILIATION == "D")) %>%
  mutate(dem.prop.of.reg = dem_voters / (dem_voters+rep_voters)) %>% 
  mutate(rep.prop.of.reg = rep_voters / (dem_voters+rep_voters)) %>% 
  mutate(unaffiliated.prop = un_voters / total_voters)

voters<-voters.sos %>%
  mutate(area = coalesce(TOWNSHIP,VILLAGE,WARD,CITY)) %>% 
  group_by(area) %>% 
  mutate(PARTY_AFFILIATION = replace_na(PARTY_AFFILIATION, 'U')) %>% 
  summarize(total_voters = n(),
            un_voters = sum(PARTY_AFFILIATION == "U"), 
            rep_voters = sum(PARTY_AFFILIATION == "R"), 
            dem_voters = sum(PARTY_AFFILIATION == "D")) %>%
  mutate(dem.prop.of.aff = dem_voters / (dem_voters+rep_voters)) %>% 
  mutate(rep.prop.of.aff = rep_voters / (dem_voters+rep_voters)) %>% 
  mutate(unaffiliated.prop = un_voters / total_voters)

mapANDvoters2024 <-
  left_join(map2024, voter.summary, by = c("PRC_NAME" = "PRECINCT_NAME"))

top_dem <- voters %>%
  arrange(desc(dem.prop.of.aff)) %>%
  slice_head(n = 5)

top_rep <- voters %>%
  arrange(desc(rep.prop.of.aff)) %>%
  slice_head(n = 6)

mapANDvoters2024 %>% 
  ggplot(aes(fill=dem.prop.of.reg)) +
  geom_sf()+
  labs(title = "2024's Party Affiliation(Democrats and Republicans)", 
       subtitle = "",
       fill = "", 
       caption = "Top Democratic/Republican Areas:\n\n Democratic        |        Republican\n
       LINCOLN HEIGHTS VILLAGE:97.8%         \tHARRISON TWP:88.4%\n
       CINTI WARD 18: 93.8%          \tCROSBY TWP87.9%:    \n
       CINTI WARD 7: 93.1%       \tMIAMI TWP:87.7%   \n
       CINTI WARD 17:88.8%         \tWHITEWATER TWP:87.3%\n
       CINTI WARD 13:88.4%        \tTHE VILLAGE OF INDIAN HILL CITY:81.5%\n
       ")+
  scale_fill_gradientn(colours=brewer.pal(n=10,name="Spectral"),na.value = "transparent",
                       breaks=c(0,.25,0.5,.75,1),labels=c("100%","75%","50%","75%","100%"),
                       limits=c(0,1))
```

# Turnout Prediction

First, I looked at the historical data of the percentage of turnout in Hamilton County. I compared different elections like the Government/Senator, Off-Cycle and Presidential Elections.

```{r}
turnout %>% 
  ggplot(aes(x = Year, y = Turnout, color = Type)) +
  geom_line(size = 1.2) +          # Line plot with thicker lines
  geom_point(size = 3) +           # Add points to the plot
  labs(
    title = "Voter Turnout in Hamilton County",
    x = "Year",
    y = "Turnout",
    color = "Election Type"
  ) +
  theme_minimal() +                # A clean theme
  scale_y_continuous(labels = scales::percent_format()) +  # Turnout as percentage
  scale_x_continuous(breaks = seq(2012, 2023, by = 1))    # Year ticks

```

In order to predict the voter turnout for 2024, I computed a linear regression of the presidential elections. I then combined my results with the turnout data I had to create a regression line on the graph to determine the predicted turnout for 2024's elections.

```{r}
presidential_turnout <- turnout %>%
  filter(Type == "Presidential")
model1 <- lm(Turnout ~ Year, data = presidential_turnout)

registered_2024 <- 604049

# Create a new data frame for prediction
year_2024 <- data.frame(Year = 2024)

# Predict turnout for 2024
predicted_turnout <- predict(model1, newdata = year_2024)

# The predicted number of voters
predicted_number_of_voters <- (predicted_turnout / 100) * registered_2024

# Calculated predicted turnout percentage
predicted_turnout_percentage <- (predicted_number_of_voters / registered_2024) * 100

# Create a new row for the predicted turnout
new_row <- data.frame(Year = 2024, Registered = registered_2024, Turnout = predicted_turnout_percentage, Type = "Presidential")

# Combine the historical data with the prediction
predictions <- bind_rows(turnout, new_row)

# Create the plot
final= ggplot(predictions, aes(x = Year, y = Turnout, color = Type)) +
  geom_point(size = 3) +           
  geom_line(data = filter(predictions, Year < 2024), aes(x = Year, y = Turnout, color = Type), size = 1.2)+
  geom_smooth(data = filter(predictions, Type == "Presidential"), method = "lm", aes(color = "Regression Line"), se = FALSE) +
  geom_text(data = new_row, aes(x = Year, y = Turnout, label = paste0(round(Turnout,2 ), "%")), 
            vjust = 1.3, hjust = 0.8) +
  labs(
    title = "Voter Turnout in Hamilton County",
    x = "Year",
    y = "Turnout Rate (%)",
    color = "Election Type"
  ) +
  theme_minimal() +                
  scale_y_continuous(labels = scales::percent_format()) +  # Turnout as percentage
  scale_x_continuous(breaks = seq(2012, 2024, by = 1))  

print(final)
```
