

```{r setup, message=FALSE,warning=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(tidycensus)
library(tidyverse)

census_api_key("97af98b509aba3ed68aa855dafc64224127cfa33")
```

# Daniel and Eden - Anna Alba


## Original graphs
```{r}
cohort_names <- c("2 Person", "3 Person", "4 Person", "5 Person",
                  "6 Person", "7 or more")

family_house <- 3:8 %>%
  sprintf("B11016_%03d", .) %>%
  set_names(cohort_names)

non_family_home <- 11:16 %>%
  sprintf("B11016_%03d", .) %>%
  set_names(cohort_names)

single_home <- "B11016_010"


family_cinci <- get_acs(
  geography = "county",
  variables = family_house,
  state = "OH",
  county = "Hamilton",
  year = 2021
) %>% 
  mutate(Status = "Family")

non_family_cinci <- get_acs(
  geography = "county",
  variables = non_family_home,
  state = "OH",
  county = "Hamilton",
  year = 2021
) %>% 
  mutate(Status = "Not Family")

single_cinci <- get_acs(
  geography = "county",
  variables = single_home,
  state = "OH",
  county = "Hamilton",
  year = 2021
) %>% 
  mutate(Status = "Not Family")

housing_data <- bind_rows(family_cinci, non_family_cinci, single_cinci)


ggplot(data = housing_data, aes(x = variable, y = estimate, fill = Status)) +
  geom_col() +
  scale_x_discrete(labels = c("2 Person", "3 Person", "4 Person", "5 Person", "6 Person", "7 or More", "1 Person")) +
  labs(x = "House Size", 
       y = "Count", 
       title = "Cincinnati Household Sizes",
       fill = "House Type")
```


## Basic ideas

The point you are trying to illustrate here is that Cincinnati is a "family city".  

* One thing jumping out at me is the fact that the ``1 person'' bar is out on the right.  Shouldn't it be on the left? 
* Also "Count" is almost never a great label.
* Last thing I have here is that maybe I'm not convinced Cincinnati really is a "family city." It seems there are lot of people alone, not with family.  What would convince me?  Maybe the same graphic for Columbus, Louisville, Cleveland,.. ?  Showing me that Cincinnati is in fact more "family-oriented" than peer cities.

```{r}
cohort_names <- c("2 Person", "3 Person", "4 Person", "5 Person",
                  "6 Person", "7 or more")

family_house <- 3:8 %>%
  sprintf("B11016_%03d", .) %>%
  set_names(cohort_names)

non_family_home <- 11:16 %>%
  sprintf("B11016_%03d", .) %>%
  set_names(cohort_names)

single_home <- "B11016_010" %>% 
  set_names("1 Person")

family_cinci <- get_acs(
  geography = "county",
  variables = family_house,
  state = "OH",
  county = "Hamilton",
  year = 2021
) %>% 
  mutate(Status = "Family")

non_family_cinci <- get_acs(
  geography = "county",
  variables = non_family_home,
  state = "OH",
  county = "Hamilton",
  year = 2021
) %>% 
  mutate(Status = "Not Family")

single_cinci <- get_acs(
  geography = "county",
  variables = single_home,
  state = "OH",
  county = "Hamilton",
  year = 2021
) %>% 
  mutate(Status = "Not Family")

housing_data <- bind_rows(family_cinci, non_family_cinci, single_cinci)


ggplot(data = housing_data, aes(x = variable, y = estimate, fill = Status)) +
  geom_col() +
  scale_x_discrete(labels = c("1 Person","2 Person", "3 Person", "4 Person", "5 Person", "6 Person", "7 or More")) +
  labs(x = "House Size", 
       y = "", 
       title = "Cincinnati Household Sizes",
       fill = "House Type")
```

# Doyinsola, Olivia and Jack - Reggie Harris

Graphs to show occupations of residents of Cincinnati.  Not sure how occupations in particular tie in directly with Reggie Harris' campaign.  Also, I am concerned about the numbers in this graph.  "Protective Law" in particular has 1.2 million individuals in it. I don't think 1.2 million people live in Cincinnati overall.

## Original graphic
```{r}
occupation_data <- get_acs(
geography = "county subdivision",
variables = c(
service = "B24011_018",
healthcare = "B24011_019",
protective_fire = "B24011_021",
protective_law = "B24011_022",
food_prep = "B24011_023",
building_maintenance = "B24011_024",
personal_care = "B24011_025"
),
state = "Ohio",
county = "Hamilton",
year = 2021
)

occupation_colors <- c("#FF9999", "#66CCCC", "#FFCC99", "#99CC66", "#FF9966", "#CC99CC", "#99CC99")

ggplot(occupation_data, aes(x = reorder(variable, estimate), y = estimate, fill = variable)) +
geom_col() +
scale_fill_manual(values = occupation_colors)+
coord_flip() +
labs(
x = "Occupation",
y = "Number of Individuals",
title = "Service Occupations in Cincinnati",
fill = "",
caption = "Data source: 2021 ACS & tidycensus R package"
) +
theme_bw()
```
## Graph changes

Two things jump out at me:

* the labels in the legend are a bit redundant since the bars are labeled.
* I don't think we want underscores in our presentation graphics.

```{r}
occupation_data <- get_acs(
geography = "county subdivision",
variables = c(
service = "B24011_018",
healthcare = "B24011_019",
`protective fire` = "B24011_021",
`protective law` = "B24011_022",
`food prep` = "B24011_023",
`building maintenance` = "B24011_024",
`personal care` = "B24011_025"
),
state = "Ohio",
county = "Hamilton",
year = 2021
)

occupation_colors <- c("#FF9999", "#66CCCC", "#FFCC99", "#99CC66", "#FF9966", "#CC99CC", "#99CC99")

ggplot(occupation_data, aes(x = variable, y = estimate, fill = variable)) +
geom_col() +
#scale_fill_manual(values = occupation_colors)+
coord_flip() +
labs(
x = "Occupation",
y = "Number of Individuals",
title = "Service Occupations in Cincinnati",
#fill = "",
caption = "Data source: 2021 ACS & tidycensus R package"
) +
theme_bw()+
  theme(legend.position="none")
```


# Aiden and Matt - Liz Keating

Graphs to show the rise in rent cost in Cincinnati.

## Original graphs
```{r }
price_names <- c("Housing < $100",
                 "Housing $100 - $199",
                 "Housing $200 - $299",
                 "Housing $300 - $399",
                 "Housing $400 - $499",
                 "Housing $500 - $599",
                 "Housing $600 - $699",
                 "Housing $700 - $799",
                 "Housing $800 - $899",
                 "Housing $900 - $999",
                 "Housing $1000 - $1499",
                 "Housing $1500 - $1999",
                 "Housing $2000 - $2499",
                 "Housing $2500 - $2999",
                 "Housing > $3000")

poo2021 <- get_acs(geography = "county",
               state = "OH",
               county = "Hamilton",
               variables = c("Housing < $100" = "B25104_002",
                             "Housing $100 - $199" = "B25104_003",
                             "Housing $200 - $299" = "B25104_004",
                             "Housing $300 - $399" = "B25104_005",
                             "Housing $400 - $499" = "B25104_006",
                             "Housing $500 - $599" = "B25104_007",
                             "Housing $600 - $699" = "B25104_008",
                             "Housing $700 - $799" = "B25104_009",
                             "Housing $800 - $899" = "B25104_010",
                             "Housing $900 - $999" = "B25104_011",
                             "Housing $1000 - $1499" = "B25104_012",
                             "Housing $1500 - $1999" = "B25104_013",
                             "Housing $2000 - $2499" = "B25104_014",
                             "Housing $2500 - $2999" = "B25104_015",
                             "Housing > $3000" = "B25104_016"
               ),
               year = 2021,
               sumfile = "acs1")

poo2020 <- get_acs(geography = "county",
               state = "OH",
               county = "Hamilton",
               variables = c("Housing < $100" = "B25104_002",
                             "Housing $100 - $199" = "B25104_003",
                             "Housing $200 - $299" = "B25104_004",
                             "Housing $300 - $399" = "B25104_005",
                             "Housing $400 - $499" = "B25104_006",
                             "Housing $500 - $599" = "B25104_007",
                             "Housing $600 - $699" = "B25104_008",
                             "Housing $700 - $799" = "B25104_009",
                             "Housing $800 - $899" = "B25104_010",
                             "Housing $900 - $999" = "B25104_011",
                             "Housing $1000 - $1499" = "B25104_012",
                             "Housing $1500 - $1999" = "B25104_013",
                             "Housing $2000 - $2499" = "B25104_014",
                             "Housing $2500 - $2999" = "B25104_015",
                             "Housing > $3000" = "B25104_016"
               ),
               year = 2020,
               sumfile = "acs1")

poo2019 <- get_acs(geography = "county",
               state = "OH",
               county = "Hamilton",
               variables = c("Housing < $100" = "B25104_002",
                             "Housing $100 - $199" = "B25104_003",
                             "Housing $200 - $299" = "B25104_004",
                             "Housing $300 - $399" = "B25104_005",
                             "Housing $400 - $499" = "B25104_006",
                             "Housing $500 - $599" = "B25104_007",
                             "Housing $600 - $699" = "B25104_008",
                             "Housing $700 - $799" = "B25104_009",
                             "Housing $800 - $899" = "B25104_010",
                             "Housing $900 - $999" = "B25104_011",
                             "Housing $1000 - $1499" = "B25104_012",
                             "Housing $1500 - $1999" = "B25104_013",
                             "Housing $2000 - $2499" = "B25104_014",
                             "Housing $2500 - $2999" = "B25104_015",
                             "Housing > $3000" = "B25104_016"
               ),
               year = 2019,
               sumfile = "acs1")

poo2019 %>% 
      mutate(variable = factor(variable, levels = price_names)) %>%
      ggplot(aes(x=variable,y=estimate))+
      geom_col() +
      coord_flip()+
      labs(x = "Rent Price", 
       y = "Number of Individuals", 
       title = "Housing Prices in Cincinnati - 2019") +
      geom_text(aes(label = estimate), 
            hjust = -0.2, 
            vjust = 0.5, 
            color = "black", 
            size = 3)

poo2020 %>% 
      mutate(variable = factor(variable, levels = price_names)) %>%
      ggplot(aes(x=variable,y=estimate))+
      geom_col() +
      coord_flip()+
      labs(x = "Rent Price", 
       y = "Number of Individuals", 
       title = "Housing Prices in Cincinnati - 2020") +
      geom_text(aes(label = estimate), 
            hjust = -0.2, 
            vjust = 0.5, 
            color = "black", 
            size = 3)

poo2021 %>% 
      mutate(variable = factor(variable, levels = price_names)) %>%
      ggplot(aes(x=variable,y=estimate))+
      geom_col() +
      ylim(0, 110000)+
      coord_flip()+
      labs(x = "Rent Price", 
       y = "Number of Individuals", 
       title = "Housing Prices in Cincinnati - 2021") +
        geom_text(aes(label = estimate), 
            hjust = -0.2, 
            vjust = 0.5, 
            color = "black", 
            size = 3)



       

```




## Quick ideas

You can adjust the dimensions of your figures in the code chunk statement. This can be helpful if you have a certain size space that you want your graph to fit in on a slide. 

```{r, fig.width=10,fig.height=11}
price_names <- c("Housing < $100",
                 "Housing $100 - $199",
                 "Housing $200 - $299",
                 "Housing $300 - $399",
                 "Housing $400 - $499",
                 "Housing $500 - $599",
                 "Housing $600 - $699",
                 "Housing $700 - $799",
                 "Housing $800 - $899",
                 "Housing $900 - $999",
                 "Housing $1000 - $1499",
                 "Housing $1500 - $1999",
                 "Housing $2000 - $2499",
                 "Housing $2500 - $2999",
                 "Housing > $3000")

poo2021 <- get_acs(geography = "county",
               state = "OH",
               county = "Hamilton",
               variables = c("Housing < $100" = "B25104_002",
                             "Housing $100 - $199" = "B25104_003",
                             "Housing $200 - $299" = "B25104_004",
                             "Housing $300 - $399" = "B25104_005",
                             "Housing $400 - $499" = "B25104_006",
                             "Housing $500 - $599" = "B25104_007",
                             "Housing $600 - $699" = "B25104_008",
                             "Housing $700 - $799" = "B25104_009",
                             "Housing $800 - $899" = "B25104_010",
                             "Housing $900 - $999" = "B25104_011",
                             "Housing $1000 - $1499" = "B25104_012",
                             "Housing $1500 - $1999" = "B25104_013",
                             "Housing $2000 - $2499" = "B25104_014",
                             "Housing $2500 - $2999" = "B25104_015",
                             "Housing > $3000" = "B25104_016"
               ),
               year = 2021,
               sumfile = "acs1")

poo2020 <- get_acs(geography = "county",
               state = "OH",
               county = "Hamilton",
               variables = c("Housing < $100" = "B25104_002",
                             "Housing $100 - $199" = "B25104_003",
                             "Housing $200 - $299" = "B25104_004",
                             "Housing $300 - $399" = "B25104_005",
                             "Housing $400 - $499" = "B25104_006",
                             "Housing $500 - $599" = "B25104_007",
                             "Housing $600 - $699" = "B25104_008",
                             "Housing $700 - $799" = "B25104_009",
                             "Housing $800 - $899" = "B25104_010",
                             "Housing $900 - $999" = "B25104_011",
                             "Housing $1000 - $1499" = "B25104_012",
                             "Housing $1500 - $1999" = "B25104_013",
                             "Housing $2000 - $2499" = "B25104_014",
                             "Housing $2500 - $2999" = "B25104_015",
                             "Housing > $3000" = "B25104_016"
               ),
               year = 2020,
               sumfile = "acs1")

poo2019 <- get_acs(geography = "county",
               state = "OH",
               county = "Hamilton",
               variables = c("Housing < $100" = "B25104_002",
                             "Housing $100 - $199" = "B25104_003",
                             "Housing $200 - $299" = "B25104_004",
                             "Housing $300 - $399" = "B25104_005",
                             "Housing $400 - $499" = "B25104_006",
                             "Housing $500 - $599" = "B25104_007",
                             "Housing $600 - $699" = "B25104_008",
                             "Housing $700 - $799" = "B25104_009",
                             "Housing $800 - $899" = "B25104_010",
                             "Housing $900 - $999" = "B25104_011",
                             "Housing $1000 - $1499" = "B25104_012",
                             "Housing $1500 - $1999" = "B25104_013",
                             "Housing $2000 - $2499" = "B25104_014",
                             "Housing $2500 - $2999" = "B25104_015",
                             "Housing > $3000" = "B25104_016"
               ),
               year = 2019,
               sumfile = "acs1")

poo2019 %>% 
      mutate(variable = factor(variable, levels = price_names)) %>%
      ggplot(aes(x=variable,y=estimate))+
      geom_col() +
      coord_flip()+
      labs(x = "Rent Price", 
       y = "Number of Individuals", 
       title = "Housing Prices in Cincinnati - 2019") +
      geom_text(aes(label = estimate), 
            hjust = -0.2, 
            vjust = 0.5, 
            color = "black", 
            size = 3)

poo2020 %>% 
      mutate(variable = factor(variable, levels = price_names)) %>%
      ggplot(aes(x=variable,y=estimate))+
      geom_col() +
      coord_flip()+
      labs(x = "Rent Price", 
       y = "Number of Individuals", 
       title = "Housing Prices in Cincinnati - 2020") +
      geom_text(aes(label = estimate), 
            hjust = -0.2, 
            vjust = 0.5, 
            color = "black", 
            size = 3)

poo2021 %>% 
      mutate(variable = factor(variable, levels = price_names)) %>%
      ggplot(aes(x=variable,y=estimate))+
      geom_col() +
      coord_flip()+
      labs(x = "Rent Price", 
       y = "Number of Individuals", 
       title = "Housing Prices in Cincinnati - 2021") +
        geom_text(aes(label = estimate), 
            hjust = -0.2, 
            vjust = 0.5, 
            color = "black", 
            size = 3)



       

```

It is probably more helpful when you make the dimensions relatively small. This has the effect of making fonts, graphics bigger and can improve clarity in the image that you actually put into a presentation.

```{r, fig.width=1.5,fig.height=2.5}
price_names <- c("Housing < $100",
                 "Housing $100 - $199",
                 "Housing $200 - $299",
                 "Housing $300 - $399",
                 "Housing $400 - $499",
                 "Housing $500 - $599",
                 "Housing $600 - $699",
                 "Housing $700 - $799",
                 "Housing $800 - $899",
                 "Housing $900 - $999",
                 "Housing $1000 - $1499",
                 "Housing $1500 - $1999",
                 "Housing $2000 - $2499",
                 "Housing $2500 - $2999",
                 "Housing > $3000")

poo2021 <- get_acs(geography = "county",
               state = "OH",
               county = "Hamilton",
               variables = c("Housing < $100" = "B25104_002",
                             "Housing $100 - $199" = "B25104_003",
                             "Housing $200 - $299" = "B25104_004",
                             "Housing $300 - $399" = "B25104_005",
                             "Housing $400 - $499" = "B25104_006",
                             "Housing $500 - $599" = "B25104_007",
                             "Housing $600 - $699" = "B25104_008",
                             "Housing $700 - $799" = "B25104_009",
                             "Housing $800 - $899" = "B25104_010",
                             "Housing $900 - $999" = "B25104_011",
                             "Housing $1000 - $1499" = "B25104_012",
                             "Housing $1500 - $1999" = "B25104_013",
                             "Housing $2000 - $2499" = "B25104_014",
                             "Housing $2500 - $2999" = "B25104_015",
                             "Housing > $3000" = "B25104_016"
               ),
               year = 2021,
               sumfile = "acs1")

poo2020 <- get_acs(geography = "county",
               state = "OH",
               county = "Hamilton",
               variables = c("Housing < $100" = "B25104_002",
                             "Housing $100 - $199" = "B25104_003",
                             "Housing $200 - $299" = "B25104_004",
                             "Housing $300 - $399" = "B25104_005",
                             "Housing $400 - $499" = "B25104_006",
                             "Housing $500 - $599" = "B25104_007",
                             "Housing $600 - $699" = "B25104_008",
                             "Housing $700 - $799" = "B25104_009",
                             "Housing $800 - $899" = "B25104_010",
                             "Housing $900 - $999" = "B25104_011",
                             "Housing $1000 - $1499" = "B25104_012",
                             "Housing $1500 - $1999" = "B25104_013",
                             "Housing $2000 - $2499" = "B25104_014",
                             "Housing $2500 - $2999" = "B25104_015",
                             "Housing > $3000" = "B25104_016"
               ),
               year = 2020,
               sumfile = "acs1")

poo2019 <- get_acs(geography = "county",
               state = "OH",
               county = "Hamilton",
               variables = c("Housing < $100" = "B25104_002",
                             "Housing $100 - $199" = "B25104_003",
                             "Housing $200 - $299" = "B25104_004",
                             "Housing $300 - $399" = "B25104_005",
                             "Housing $400 - $499" = "B25104_006",
                             "Housing $500 - $599" = "B25104_007",
                             "Housing $600 - $699" = "B25104_008",
                             "Housing $700 - $799" = "B25104_009",
                             "Housing $800 - $899" = "B25104_010",
                             "Housing $900 - $999" = "B25104_011",
                             "Housing $1000 - $1499" = "B25104_012",
                             "Housing $1500 - $1999" = "B25104_013",
                             "Housing $2000 - $2499" = "B25104_014",
                             "Housing $2500 - $2999" = "B25104_015",
                             "Housing > $3000" = "B25104_016"
               ),
               year = 2019,
               sumfile = "acs1")

poo2019 %>% 
      mutate(variable = factor(variable, levels = price_names)) %>%
      ggplot(aes(x=variable,y=estimate))+
      geom_col() +
      coord_flip()+
      labs(x = "Rent Price", 
       y = "Number of Individuals", 
       title = "Housing Prices in Cincinnati - 2019") +
      geom_text(aes(label = estimate), 
            hjust = -0.2, 
            vjust = 0.5, 
            color = "black", 
            size = 3)

poo2020 %>% 
      mutate(variable = factor(variable, levels = price_names)) %>%
      ggplot(aes(x=variable,y=estimate))+
      geom_col() +
      coord_flip()+
      labs(x = "Rent Price", 
       y = "Number of Individuals", 
       title = "Housing Prices in Cincinnati - 2020") +
      geom_text(aes(label = estimate), 
            hjust = -0.2, 
            vjust = 0.5, 
            color = "black", 
            size = 3)

poo2021 %>% 
      mutate(variable = factor(variable, levels = price_names)) %>%
      ggplot(aes(x=variable,y=estimate))+
      geom_col() +
      coord_flip()+
      labs(x = "Rent Price", 
       y = "Number of Individuals", 
       title = "Housing Prices in Cincinnati - 2021") +
        geom_text(aes(label = estimate), 
            hjust = -0.2, 
            vjust = 0.5, 
            color = "black", 
            size = 3)

```

Lose the word "Housing in front of every label on the vertical axis and also change "Housing" to "Rent" in the title

```{r, fig.width=1.5,fig.height=2.5}
price_names <- c("< $100",
                 "$100 - $199",
                 "$200 - $299",
                 "$300 - $399",
                 "$400 - $499",
                 "$500 - $599",
                 "$600 - $699",
                 "$700 - $799",
                 "$800 - $899",
                 "$900 - $999",
                 "$1000 - $1499",
                 "$1500 - $1999",
                 "$2000 - $2499",
                 "$2500 - $2999",
                 "> $3000")

poo2021 <- get_acs(geography = "county",
               state = "OH",
               county = "Hamilton",
               variables = c("< $100" = "B25104_002",
                             "$100 - $199" = "B25104_003",
                             "$200 - $299" = "B25104_004",
                             "$300 - $399" = "B25104_005",
                             "$400 - $499" = "B25104_006",
                             "$500 - $599" = "B25104_007",
                             "$600 - $699" = "B25104_008",
                             "$700 - $799" = "B25104_009",
                             "$800 - $899" = "B25104_010",
                             "$900 - $999" = "B25104_011",
                             "$1000 - $1499" = "B25104_012",
                             "$1500 - $1999" = "B25104_013",
                             "$2000 - $2499" = "B25104_014",
                             "$2500 - $2999" = "B25104_015",
                             "> $3000" = "B25104_016"
               ),
               year = 2021,
               sumfile = "acs1")

poo2020 <- get_acs(geography = "county",
               state = "OH",
               county = "Hamilton",
               variables = c("< $100" = "B25104_002",
                             "$100 - $199" = "B25104_003",
                             "$200 - $299" = "B25104_004",
                             "$300 - $399" = "B25104_005",
                             "$400 - $499" = "B25104_006",
                             "$500 - $599" = "B25104_007",
                             "$600 - $699" = "B25104_008",
                             "$700 - $799" = "B25104_009",
                             "$800 - $899" = "B25104_010",
                             "$900 - $999" = "B25104_011",
                             "$1000 - $1499" = "B25104_012",
                             "$1500 - $1999" = "B25104_013",
                             "$2000 - $2499" = "B25104_014",
                             "$2500 - $2999" = "B25104_015",
                             "> $3000" = "B25104_016"
               ),
               year = 2020,
               sumfile = "acs1")

poo2019 <- get_acs(geography = "county",
               state = "OH",
               county = "Hamilton",
               variables = c("< $100" = "B25104_002",
                             "$100 - $199" = "B25104_003",
                             "$200 - $299" = "B25104_004",
                             "$300 - $399" = "B25104_005",
                             "$400 - $499" = "B25104_006",
                             "$500 - $599" = "B25104_007",
                             "$600 - $699" = "B25104_008",
                             "$700 - $799" = "B25104_009",
                             "$800 - $899" = "B25104_010",
                             "$900 - $999" = "B25104_011",
                             "$1000 - $1499" = "B25104_012",
                             "$1500 - $1999" = "B25104_013",
                             "$2000 - $2499" = "B25104_014",
                             "$2500 - $2999" = "B25104_015",
                             "> $3000" = "B25104_016"
               ),
               year = 2019,
               sumfile = "acs1")

poo2019 %>% 
      mutate(variable = factor(variable, levels = price_names)) %>%
      ggplot(aes(x=variable,y=estimate))+
      geom_col() +
      coord_flip()+
      labs(x = "Rent Price", 
       y = "Number of Individuals", 
       title = "Rent Prices in Cincinnati - 2019") +
      geom_text(aes(label = estimate), 
            hjust = -0.2, 
            vjust = 0.5, 
            color = "black", 
            size = 3)

poo2020 %>% 
      mutate(variable = factor(variable, levels = price_names)) %>%
      ggplot(aes(x=variable,y=estimate))+
      geom_col() +
      coord_flip()+
      labs(x = "Rent Price", 
       y = "Number of Individuals", 
       title = "Rent Prices in Cincinnati - 2020") +
      geom_text(aes(label = estimate), 
            hjust = -0.2, 
            vjust = 0.5, 
            color = "black", 
            size = 3)

poo2021 %>% 
      mutate(variable = factor(variable, levels = price_names)) %>%
      ggplot(aes(x=variable,y=estimate))+
      geom_col() +
      coord_flip()+
      labs(x = "Rent Price", 
       y = "Number of Individuals", 
       title = "Rent Prices in Cincinnati - 2021") +
        geom_text(aes(label = estimate), 
            hjust = -0.2, 
            vjust = 0.5, 
            color = "black", 
            size = 3)

```


I'm not sure we need to label the numbers for the barplots

```{r, fig.width=1.5,fig.height=2.5}
price_names <- c("< $100",
                 "$100 - $199",
                 "$200 - $299",
                 "$300 - $399",
                 "$400 - $499",
                 "$500 - $599",
                 "$600 - $699",
                 "$700 - $799",
                 "$800 - $899",
                 "$900 - $999",
                 "$1000 - $1499",
                 "$1500 - $1999",
                 "$2000 - $2499",
                 "$2500 - $2999",
                 "> $3000")

poo2021 <- get_acs(geography = "county",
               state = "OH",
               county = "Hamilton",
               variables = c("< $100" = "B25104_002",
                             "$100 - $199" = "B25104_003",
                             "$200 - $299" = "B25104_004",
                             "$300 - $399" = "B25104_005",
                             "$400 - $499" = "B25104_006",
                             "$500 - $599" = "B25104_007",
                             "$600 - $699" = "B25104_008",
                             "$700 - $799" = "B25104_009",
                             "$800 - $899" = "B25104_010",
                             "$900 - $999" = "B25104_011",
                             "$1000 - $1499" = "B25104_012",
                             "$1500 - $1999" = "B25104_013",
                             "$2000 - $2499" = "B25104_014",
                             "$2500 - $2999" = "B25104_015",
                             "> $3000" = "B25104_016"
               ),
               year = 2021,
               sumfile = "acs1")

poo2020 <- get_acs(geography = "county",
               state = "OH",
               county = "Hamilton",
               variables = c("< $100" = "B25104_002",
                             "$100 - $199" = "B25104_003",
                             "$200 - $299" = "B25104_004",
                             "$300 - $399" = "B25104_005",
                             "$400 - $499" = "B25104_006",
                             "$500 - $599" = "B25104_007",
                             "$600 - $699" = "B25104_008",
                             "$700 - $799" = "B25104_009",
                             "$800 - $899" = "B25104_010",
                             "$900 - $999" = "B25104_011",
                             "$1000 - $1499" = "B25104_012",
                             "$1500 - $1999" = "B25104_013",
                             "$2000 - $2499" = "B25104_014",
                             "$2500 - $2999" = "B25104_015",
                             "> $3000" = "B25104_016"
               ),
               year = 2020,
               sumfile = "acs1")

poo2019 <- get_acs(geography = "county",
               state = "OH",
               county = "Hamilton",
               variables = c("< $100" = "B25104_002",
                             "$100 - $199" = "B25104_003",
                             "$200 - $299" = "B25104_004",
                             "$300 - $399" = "B25104_005",
                             "$400 - $499" = "B25104_006",
                             "$500 - $599" = "B25104_007",
                             "$600 - $699" = "B25104_008",
                             "$700 - $799" = "B25104_009",
                             "$800 - $899" = "B25104_010",
                             "$900 - $999" = "B25104_011",
                             "$1000 - $1499" = "B25104_012",
                             "$1500 - $1999" = "B25104_013",
                             "$2000 - $2499" = "B25104_014",
                             "$2500 - $2999" = "B25104_015",
                             "> $3000" = "B25104_016"
               ),
               year = 2019,
               sumfile = "acs1")

poo2019 %>% 
      mutate(variable = factor(variable, levels = price_names)) %>%
      ggplot(aes(x=variable,y=estimate))+
      geom_col() +
      coord_flip()+
      labs(x = "Rent Price", 
       y = "Number of Individuals", 
       title = "Rent Prices in Cincinnati - 2019") 
# +
      # geom_text(aes(label = estimate), 
      #       hjust = -0.2, 
      #       vjust = 0.5, 
      #       color = "black", 
      #       size = 3)

poo2020 %>% 
      mutate(variable = factor(variable, levels = price_names)) %>%
      ggplot(aes(x=variable,y=estimate))+
      geom_col() +
      coord_flip()+
      labs(x = "Rent Price", 
       y = "Number of Individuals", 
       title = "Rent Prices in Cincinnati - 2020") 
# +
      # geom_text(aes(label = estimate), 
      #       hjust = -0.2, 
      #       vjust = 0.5, 
      #       color = "black", 
      #       size = 3)

poo2021 %>% 
      mutate(variable = factor(variable, levels = price_names)) %>%
      ggplot(aes(x=variable,y=estimate))+
      geom_col() +
      coord_flip()+
      labs(x = "Rent Price", 
       y = "Number of Individuals", 
       title = "Rent Prices in Cincinnati - 2021") 
# +
        # geom_text(aes(label = estimate), 
        #     hjust = -0.2, 
        #     vjust = 0.5, 
        #     color = "black", 
        #     size = 3)

```


## Bigger ideas

The point of our graphics is to clearly demonstrate the point that we want to make.  In this case, that's the rent prices have increased. This should make us think hard about what would be the clearest visualization of that increase.

My first idea for this is to try to get things on a single graph, rather than needing to compare two separate graphs and point out the difference.

Option 1:

* put 2019 and 2021 data into one dataframe and then make single graph with both in it.
* you can also change font size directly without changing the size of your window.  In this example, we make font size bigger for the whole graphic but you can do this individually (i.e. axes labels, title, etc) too if you prefer.


```{r}
price_names <- c("< $100",
                 "$100 - $199",
                 "$200 - $299",
                 "$300 - $399",
                 "$400 - $499",
                 "$500 - $599",
                 "$600 - $699",
                 "$700 - $799",
                 "$800 - $899",
                 "$900 - $999",
                 "$1000 - $1499",
                 "$1500 - $1999",
                 "$2000 - $2499",
                 "$2500 - $2999",
                 "> $3000")

poo2021 <- get_acs(geography = "county",
               state = "OH",
               county = "Hamilton",
               variables = c("< $100" = "B25104_002",
                             "$100 - $199" = "B25104_003",
                             "$200 - $299" = "B25104_004",
                             "$300 - $399" = "B25104_005",
                             "$400 - $499" = "B25104_006",
                             "$500 - $599" = "B25104_007",
                             "$600 - $699" = "B25104_008",
                             "$700 - $799" = "B25104_009",
                             "$800 - $899" = "B25104_010",
                             "$900 - $999" = "B25104_011",
                             "$1000 - $1499" = "B25104_012",
                             "$1500 - $1999" = "B25104_013",
                             "$2000 - $2499" = "B25104_014",
                             "$2500 - $2999" = "B25104_015",
                             "> $3000" = "B25104_016"
               ),
               year = 2021,
               sumfile = "acs1")

poo2020 <- get_acs(geography = "county",
               state = "OH",
               county = "Hamilton",
               variables = c("< $100" = "B25104_002",
                             "$100 - $199" = "B25104_003",
                             "$200 - $299" = "B25104_004",
                             "$300 - $399" = "B25104_005",
                             "$400 - $499" = "B25104_006",
                             "$500 - $599" = "B25104_007",
                             "$600 - $699" = "B25104_008",
                             "$700 - $799" = "B25104_009",
                             "$800 - $899" = "B25104_010",
                             "$900 - $999" = "B25104_011",
                             "$1000 - $1499" = "B25104_012",
                             "$1500 - $1999" = "B25104_013",
                             "$2000 - $2499" = "B25104_014",
                             "$2500 - $2999" = "B25104_015",
                             "> $3000" = "B25104_016"
               ),
               year = 2020,
               sumfile = "acs1")

poo2019 <- get_acs(geography = "county",
               state = "OH",
               county = "Hamilton",
               variables = c("< $100" = "B25104_002",
                             "$100 - $199" = "B25104_003",
                             "$200 - $299" = "B25104_004",
                             "$300 - $399" = "B25104_005",
                             "$400 - $499" = "B25104_006",
                             "$500 - $599" = "B25104_007",
                             "$600 - $699" = "B25104_008",
                             "$700 - $799" = "B25104_009",
                             "$800 - $899" = "B25104_010",
                             "$900 - $999" = "B25104_011",
                             "$1000 - $1499" = "B25104_012",
                             "$1500 - $1999" = "B25104_013",
                             "$2000 - $2499" = "B25104_014",
                             "$2500 - $2999" = "B25104_015",
                             "> $3000" = "B25104_016"
               ),
               year = 2019,
               sumfile = "acs1")

## Add in a column for which year each data point is from and put them into one dataframe
mut2019 <-
  poo2019 %>% 
  mutate(year = '2019')

mut2021 <- poo2021 %>% 
  mutate(year = '2021')

poop = bind_rows(mut2019,mut2021)


## plot them on the same axes.

poop %>% 
      mutate(variable = factor(variable, levels = price_names)) %>%
      ggplot(aes(x=variable,y=estimate,fill=year,group=year))+
      geom_col(position = "dodge") +
      coord_flip()+
      labs(x = "Rent Price", 
       y = "Number of Individuals", 
       title = "Rent Prices in Cincinnati") +
   theme(text = element_text(size = 15))           
```

The last effort was pretty solid, but let's try one more. Instead of bars, let's try to show the shift in distribution by smoothing out the data with curves for 2019 and 2021.  The problem with "smoothing" like this is that it involves regression and you can't do regression on a label like ``$100-199.''  R sees that as a name and number a number.


```{r}
price_names <- c("< $100",
                 "$100 - $199",
                 "$200 - $299",
                 "$300 - $399",
                 "$400 - $499",
                 "$500 - $599",
                 "$600 - $699",
                 "$700 - $799",
                 "$800 - $899",
                 "$900 - $999",
                 "$1000 - $1499",
                 "$1500 - $1999",
                 "$2000 - $2499",
                 "$2500 - $2999",
                 "> $3000")

price_midpoints <- c(50,
                 150,
                 250,
                 350,
                 450,
                 550,
                 650,
                 750,
                 850,
                 950,
                 1250,
                 1750,
                 2250,
                 2750,
                 3250)

poo2021 <- get_acs(geography = "county",
               state = "OH",
               county = "Hamilton",
               variables = c("< $100" = "B25104_002",
                             "$100 - $199" = "B25104_003",
                             "$200 - $299" = "B25104_004",
                             "$300 - $399" = "B25104_005",
                             "$400 - $499" = "B25104_006",
                             "$500 - $599" = "B25104_007",
                             "$600 - $699" = "B25104_008",
                             "$700 - $799" = "B25104_009",
                             "$800 - $899" = "B25104_010",
                             "$900 - $999" = "B25104_011",
                             "$1000 - $1499" = "B25104_012",
                             "$1500 - $1999" = "B25104_013",
                             "$2000 - $2499" = "B25104_014",
                             "$2500 - $2999" = "B25104_015",
                             "> $3000" = "B25104_016"
               ),
               year = 2021,
               sumfile = "acs1") %>% 
  mutate("price_midpoints"= price_midpoints)


poo2019 <- get_acs(geography = "county",
               state = "OH",
               county = "Hamilton",
               variables = c("< $100" = "B25104_002",
                             "$100 - $199" = "B25104_003",
                             "$200 - $299" = "B25104_004",
                             "$300 - $399" = "B25104_005",
                             "$400 - $499" = "B25104_006",
                             "$500 - $599" = "B25104_007",
                             "$600 - $699" = "B25104_008",
                             "$700 - $799" = "B25104_009",
                             "$800 - $899" = "B25104_010",
                             "$900 - $999" = "B25104_011",
                             "$1000 - $1499" = "B25104_012",
                             "$1500 - $1999" = "B25104_013",
                             "$2000 - $2499" = "B25104_014",
                             "$2500 - $2999" = "B25104_015",
                             "> $3000" = "B25104_016"
               ),
               year = 2019,
               sumfile = "acs1") %>% 
  mutate("price_midpoints"= price_midpoints)

## Add in a column for which year each data point is from and put them into one dataframe
mut2019 <-
  poo2019 %>% 
  mutate(year = '2019')

mut2021 <- poo2021 %>% 
  mutate(year = '2021')

poop = bind_rows(mut2019,mut2021)



poop %>% 
      ggplot(aes(x=price_midpoints,y=estimate,group=year,color=year))+
      geom_density(stat='identity',lwd=1) +
      #coord_flip()+
      labs(x = "Rent Price", 
       y = "Number of Individuals", 
       title = "Rent Prices in Cincinnati") +
   theme(text = element_text(size = 15))           
```


