---
title: "Statistics_project"
author: "Stephen Lynch & Abigail Sandquist"
format: pdf
---

## Appendix ##

  This project investigates how age and years of experience affect employee salary using a dataset containing age, gender, job title, years of experience, and salary. The analysis examines each of the research questions outlined in the proposal: the strength of the relationship between years of experience and salary, whether age contributes additional explanatory power beyond experience, whether salary patterns differ across genders or job titles, the rate at which salary increases with additional experience, and whether salary can be accurately predicted using statistical models. Using descriptive statistics, data visualizations, and linear regression models, we evaluate how these variables interact and identify the most influential predictors of earnings. The findings aim to clarify how professional growth impacts salary progression and provide insight into broader trends in compensation across demographic and job-related categories. This study contributes to understanding the key determinants of salary and supports future work in salary prediction and workforce analytics.

## Introduction ##

  Understanding what drives differences in employee salaries is an important question in both workforce analytics and applied statistics. Salary is influenced by many factors, but two of the most commonly discussed predictors are age and years of experience. As individuals gain professional experience, they typically acquire new skills, demonstrate increased responsibility, and advance within their roles—all of which are expected to contribute to higher earnings. Age may also play a role, as it often reflects career stage, career longevity, and time spent in the labor force. However, the degree to which age and experience independently or jointly influence salary remains an empirical question and can vary across industries, job titles, and demographic groups.
  This research is significant because salary growth is closely tied to career development, economic mobility, and long-term financial stability. Employers use salary models to make compensation decisions, while individuals rely on salary expectations to make choices about education, job changes, and skill development. Understanding how much experience contributes to salary, how strongly age correlates with earnings, and whether demographic or job-related disparities exist can provide valuable insight for employees, employers, policymakers, and workforce researchers.
  Despite the importance of these issues, salary data is often complex and influenced by multiple interacting variables. This project uses statistical methods to analyze a dataset containing age, gender, job title, years of experience, and salary to better understand the relationships among these variables. Specifically, we investigate how salary changes with increasing experience, whether age adds predictive value beyond experience, whether salary differs across genders or job titles, and whether statistical models can accurately predict salary based on these factors.
  By examining these relationships, this research aims to deepen our understanding of the key determinants of salary and to highlight potential differences or inequalities within the workforce. The results contribute to broader discussions about compensation fairness, career progression, and the factors that most strongly influence earnings in professional environments.


```{r}
library(tidyverse)

df <- read_csv("Salary_Data.csv")


df <- df %>% 
  mutate(
    `Education Level` = `Education Level` %>%
      str_remove(" Degree")) %>% 
  subset(`Education Level` != "NA")

df[5876,3] <- "PhD"    

```


```{r}

salary_years_exp <- lm(data = df, Salary ~ `Years of Experience`)

summary(salary_years_exp)


```
We can see there is a substansial impact on salary caused by Years of Expirience,
for every year increase in YoE the lm estimates about a $7046 increase in Salary. 
The Caveat is that we only have an R^2 of about .65 meaning that we are only explaining
about 65% of the total variance of the data. 



```{r}

df$Age_YoE <- df$Age * df$`Years of Experience`

lm2 <- lm(data = df, Salary ~ `Years of Experience` + `Age` + Age_YoE)

summary(lm2)

aov_model <- aov(Salary ~ `Years of Experience` + `Age` + Age_YoE , data = df)
summary(aov_model)

```



```{r}

lm3 <- lm(data = df, Salary ~ `Job Title`)
summary(lm3)

lm4 <- lm(data = df, Salary ~ Gender)
summary(lm4)


```



```{r}

lm5 <- lm(data = df, Salary ~ Age + Gender + `Education Level` + `Years of Experience`)
lm6 <- lm(data = df, Salary ~ Age + Gender + `Education Level` + `Years of Experience` + Age_YoE)
summary(lm5)
summary(lm6)
```
```{r}

plot(lm6, which = 1)

```



``` {r}

df %>% 
  ggplot(aes(y = Salary, x = Age))+
  geom_point()+
  facet_wrap(~`Education Level`)


```


``` {r}

df %>% 
  ggplot(aes(y = Salary, x = `Years of Experience`))+
  geom_point()+
  facet_wrap(~`Education Level`)


```




``` {r}

df %>% 
  ggplot(aes(y = Salary, x = Gender))+
  geom_boxplot()+
  facet_wrap(~`Education Level`)


```


``` {r}

df %>% 
  ggplot(aes(y = Salary, x = `Job Title`))+
  geom_point()+
  facet_wrap(~`Education Level`)


```



Make a starting LM for the data with Salary as our Dependent Variable and years
of experience and age as Independant


``` {r}

model1 <- lm(Salary ~ `Years of Experience` + log(Age) + Gender + `Years of Experience` + `Education Level` + `Job Title`, data = df)
model2 <- lm(Salary ~ `Years of Experience` + log(Age) + Gender + `Years of Experience` + `Education Level` , data = df)
summary(model2)
summary(model1)

coefs2 <- summary(model2)$coefficients
sig_coefs2 <- coefs2[coefs2[, "Pr(>|t|)"] < 0.05, ]
sig_coefs2

coefs1 <- summary(model1)$coefficients
sig_coefs1 <- coefs1[coefs1[, "Pr(>|t|)"] < 0.05, ]
sig_coefs1

final_model <- step(model1, direction = "backward")
summary(final_model)
anova(final_model)
```


```{r}

plot(model2, which = 1)

plot(model1, which = 1)

plot(final_model, which = 1)

```




```{r}
df$`Years of Experience2` = log(df$`Years of Experience`)

df$Age2 = log(df$Age)


df$salary2 = log(df$Salary, base = exp(1))



loged_model1 <- lm(salary2 ~ `Years of Experience` + Age + Gender + `Years of Experience` + `Education Level` + `Job Title`, data = df)
loged_model2 <- lm(salary2 ~ `Years of Experience` + Age + Gender + `Years of Experience` + `Education Level` , data = df)
summary(loged_model2)
summary(loged_model1)

coefs2 <- summary(loged_model2)$coefficients
sig_coefs2 <- coefs2[coefs2[, "Pr(>|t|)"] < 0.05, ]
sig_coefs2

coefs1 <- summary(loged_model1)$coefficients
sig_coefs1 <- coefs1[coefs1[, "Pr(>|t|)"] < 0.05, ]
sig_coefs1

loged_final_model <- step(loged_model1, direction = "backward")
summary(loged_final_model)
anova(loged_final_model)






```


```{r}

plot(loged_model2)

plot(loged_model1)

plot(loged_final_model)


```

## Discussion ##

  The findings of this project provide meaningful insight into how years of experience, age, gender, and job title relate to employee salary. The analysis shows that years of experience is the strongest and most consistent predictor of salary, with a clear positive association indicating that employees with more experience tend to earn higher wages. Although age also correlates with salary, the regression results suggest that much of this effect is actually captured by experience rather than age itself. This aligns with expectations that professional development, accumulated skills, and advancement opportunities are what primarily drive earnings, rather than age alone. Gender and job title also contributed to observable differences in salary outcomes. While these variables were not the central focus of the study, the patterns suggest that certain roles naturally command different salary ranges and that small differences between genders could be present depending on the composition of the dataset. These results highlight the complexity of salary determination and show that compensation is shaped not only by personal experience but also by the structural and demographic characteristics of the workforce.
  Despite the useful insights gained, the study has several limitations. The dataset may not fully represent the broader labor market, as its size, structure, or sampling method could limit generalizability. Salary is influenced by many factors not included in the dataset, such as geographic region, industry, company size, or specific skill sets, and the absence of these variables restricts the depth of the analysis. In addition, variables such as “years of experience” or “job title” can vary greatly across industries, which may introduce inconsistencies in interpretation. Because the dataset appears to be cross-sectional, the analysis identifies associations but cannot establish causation or track salary changes within individuals over time. It is also possible that certain unobserved biases exist, especially if elements of the data were self-reported or disproportionately reflect certain demographic groups or job sectors.
  The assumptions made in the analysis are generally reasonable for the context of an exploratory statistical study. The linear regression models assume a linear relationship between predictors and salary, which appears appropriate based on the visual patterns in the data, although real-world salary progression may sometimes follow nonlinear trends. The analysis assumes independence of observations, meaning that each row in the dataset represents a unique individual; if this were not the case, the conclusions could be affected. Regression diagnostics suggested that residuals were approximately normal, supporting the validity of inference, and the assumption of additivity—that each predictor contributes independently to salary—provides a useful starting point for understanding basic relationships among the variables.
  The study opens several pathways for future research. Expanding the dataset to include additional variables such as education level, geographic region, industry classification, or employer characteristics would allow for a more comprehensive model of salary determinants. A deeper investigation into gender differences or job title disparities could help determine whether certain groups face inequitable salary outcomes. Longitudinal data would strengthen future analyses by allowing researchers to observe true salary growth, career mobility, and experience accumulation over time. More advanced modeling approaches, such as machine learning algorithms or nonlinear regression techniques, could capture complex interactions that simple linear models might overlook. Finally, conducting industry-specific studies would provide more tailored insights, since salary dynamics differ widely across fields such as technology, healthcare, finance, and education.
