---
title: "Statistics_project"
author: "Stephen Lynch & Abigail Sandquist"
format: pdf
---

Read in the Data to begin assessment.


```{r}
library(tidyverse)

df <- read_csv("Salary_Data.csv")


df <- df %>% 
  mutate(
    `Education Level` = `Education Level` %>%
      str_remove(" Degree")) %>% 
  subset(`Education Level` != "NA")

df[5876,3] <- "PhD"    

```


```{r}

salary_years_exp <- lm(data = df, Salary ~ `Years of Experience`)

summary(salary_years_exp)


```
We can see there is a substansial impact on salary caused by Years of Expirience,
for every year increase in YoE the lm estimates about a $7046 increase in Salary. 
The Caveat is that we only have an R^2 of about .65 meaning that we are only explaining
about 65% of the total variance of the data. 



```{r}

df$Age_YoE <- df$Age * df$`Years of Experience`

lm2 <- lm(data = df, Salary ~ `Years of Experience` + `Age` + Age_YoE)

summary(lm2)

aov_model <- aov(Salary ~ `Years of Experience` + `Age` + Age_YoE , data = df)
summary(aov_model)

```



```{r}

lm3 <- lm(data = df, Salary ~ `Job Title`)
summary(lm3)

lm4 <- lm(data = df, Salary ~ Gender)
summary(lm4)


```



```{r}

lm5 <- lm(data = df, Salary ~ Age + Gender + `Education Level` + `Years of Experience`)
lm6 <- lm(data = df, Salary ~ Age + Gender + `Education Level` + `Years of Experience` + Age_YoE)
summary(lm5)
summary(lm6)
```
```{r}

plot(lm6, which = 1)

```



``` {r}

df %>% 
  ggplot(aes(y = Salary, x = Age))+
  geom_point()+
  facet_wrap(~`Education Level`)


```


``` {r}

df %>% 
  ggplot(aes(y = Salary, x = `Years of Experience`))+
  geom_point()+
  facet_wrap(~`Education Level`)


```




``` {r}

df %>% 
  ggplot(aes(y = Salary, x = Gender))+
  geom_boxplot()+
  facet_wrap(~`Education Level`)


```


``` {r}

df %>% 
  ggplot(aes(y = Salary, x = `Job Title`))+
  geom_point()+
  facet_wrap(~`Education Level`)


```



Make a starting LM for the data with Salary as our Dependent Variable and years
of experience and age as Independant


``` {r}

model1 <- lm(Salary ~ `Years of Experience` + log(Age) + Gender + `Years of Experience` + `Education Level` + `Job Title`, data = df)
model2 <- lm(Salary ~ `Years of Experience` + log(Age) + Gender + `Years of Experience` + `Education Level` , data = df)
summary(model2)
summary(model1)

coefs2 <- summary(model2)$coefficients
sig_coefs2 <- coefs2[coefs2[, "Pr(>|t|)"] < 0.05, ]
sig_coefs2

coefs1 <- summary(model1)$coefficients
sig_coefs1 <- coefs1[coefs1[, "Pr(>|t|)"] < 0.05, ]
sig_coefs1

final_model <- step(model1, direction = "backward")
summary(final_model)
anova(final_model)
```


```{r}

plot(model2, which = 1)

plot(model1, which = 1)

plot(final_model, which = 1)

```




```{r}
df$`Years of Experience2` = log(df$`Years of Experience`)

df$Age2 = log(df$Age)


df$salary2 = log(df$Salary, base = exp(1))



loged_model1 <- lm(salary2 ~ `Years of Experience` + Age + Gender + `Years of Experience` + `Education Level` + `Job Title`, data = df)
loged_model2 <- lm(salary2 ~ `Years of Experience` + Age + Gender + `Years of Experience` + `Education Level` , data = df)
summary(loged_model2)
summary(loged_model1)

coefs2 <- summary(loged_model2)$coefficients
sig_coefs2 <- coefs2[coefs2[, "Pr(>|t|)"] < 0.05, ]
sig_coefs2

coefs1 <- summary(loged_model1)$coefficients
sig_coefs1 <- coefs1[coefs1[, "Pr(>|t|)"] < 0.05, ]
sig_coefs1

loged_final_model <- step(loged_model1, direction = "backward")
summary(loged_final_model)
anova(loged_final_model)






```


```{r}

plot(loged_model2)

plot(loged_model1)

plot(loged_final_model)


```


