---
title: "Difference in Votes for Deters and Trump in 2020"
author: Colleen Malloy
format: html
editor: visual
---

## Looking at 2020 Election Results: Republican Votes in Presidential and Prosecutor Race (Hamilton County, Ohio)

```{r}
#| echo: false
#| results: hide
#| warning: false
#| message: false
library(readr)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(sf)
library(RColorBrewer)
library(readxl)
```

Combine 2020 precincts and 2020 election results.

Make a copy with only the columns we need, (Trump, Biden, Deters, Rucker).

Find the differences in proportion of votes and of number of votes between Republican and Democratic opponents per race.

Use these columns to create maps to visualize these results by precinct.

```{r}
#| result: hide
#| warning: false
#| message: false
results2020 <- read_excel("../data/election results/G20_Official_Canvass.xlsx", sheet = "Candidates",skip=1)
map <- st_zm(st_read("../data/maps/PRECINCT_052219.shp"))
```

```{r}

mapANDresults2020 <-
  left_join(map, results2020, by = c("PRECINCT" = "PRECINCT"))


columns <- mapANDresults2020 %>%
  select(`PRC #`, `PRECINCT`, `REGISTERED VOTERS TOTAL`, `PERCENT`,  `Biden & Harris  (Dem)`, `Trump & Pence       (Rep)`,`Joseph T. Deters      (Rep)`, `Fanon A. Rucker      (Dem)`) %>% 
  mutate(Trump.prop = `Trump & Pence       (Rep)`/( `Biden & Harris  (Dem)`+ `Trump & Pence       (Rep)`)) %>% 
  mutate(Deters.prop = `Joseph T. Deters      (Rep)`/(`Joseph T. Deters      (Rep)` + `Fanon A. Rucker      (Dem)`)) %>% 
  mutate(prop.difference.Rep = Deters.prop - Trump.prop) %>% 
  mutate(vote.difference.Rep = (`Joseph T. Deters      (Rep)`) - (`Trump & Pence       (Rep)`))

```

```{r}
library(dplyr)
library(ggplot2)
library(sf)
library(RColorBrewer)

# Coordinates for the line (adjust as needed)
start_coords <- c(-84.48, 39.02) # Example starting point
end_coords <- c(-84.515, 39.13)   # Example ending point

# Plot the map with the highlighted precinct, line, and label
columns %>%
  ggplot(aes(fill = prop.difference.Rep)) + 
  geom_sf() + 
  labs(title = "Difference in Votes for Deters and Trump in 2020 (Hamilton County, OH)",
       subtitle = "Deters won with 52.44% of votes, Trump lost with 41.28% of votes",
       fill = "Vote for Deters - Trump (%)",
       caption = "") + 
  geom_segment(aes(x = start_coords[1], y = start_coords[2], 
                   xend = end_coords[1], yend = end_coords[2]),
               arrow = arrow(length = unit(0.3, "cm")), color = "black") + 
  geom_label(aes(x = start_coords[1], y = start_coords[2], label = "University of Cincinnati"), 
             nudge_y = 0.02, color = "black", fill = "white") + 
  scale_fill_gradient2(low = "darkgreen", mid = "white", high = "purple", 
                       midpoint = 0,
                       na.value = "transparent",
                       breaks = c(-.05, 0, .05, .1, .15, .2, .22),
                       labels = c("-5", "0", "5", "10", "15", "", "22"),
                       limits = c(-.05, .22)) + 
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20),
    plot.subtitle = element_text(size = 16),
    plot.caption = element_text(size = 12),
    axis.title = element_blank(), # Removes axis titles
    axis.text = element_blank(),  # Removes axis text
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12)
  )

```

Are there any precincts that stand out?

Do the precincts with high percentages and vote counts of difference lie within our swing voters precincts?

Should these be our swing voters that we target?
