---
title: "xml stuff"
format: html
editor: visual
---

# Setup

```{r}
library(dplyr)
library(tidyr)
library(xml2)
library(tidyverse)
library(purrr)
library(sf)
library(RColorBrewer)
library(tidycensus)
census_api_key("fa67b1dbacf4fbbb1b14c875f34437c6cbdaa694")
```

Function that parses the xml results files.

```{r}
parse_contest_results <- function(xml_file){
  
  ## Get all Contest nodes
  contests <- xml_file %>% xml_find_all(".//Contest")
  
  # Function to extract vote data for each contest and candidate
  parse_contest_votes <- function(contest_node) {
    contest_name <- xml_attr(contest_node, "text")
    
    # Get all candidate (Choice) nodes
    choices <- contest_node %>% xml_find_all(".//Choice")
    
    # Iterate through each candidate
    map_df(choices, function(choice) {
      candidate_name <- xml_attr(choice, "text")
      
      # Get all VoteType nodes (Early, Election)
      vote_types <- choice %>% xml_find_all(".//VoteType")
      
      # Iterate through each VoteType and precinct data
      map_df(vote_types, function(vote_type) {
        vote_type_name <- xml_attr(vote_type, "name")
        
        # Extract precincts
        precincts <- vote_type %>% xml_find_all(".//Precinct")
        
        # For each precinct, extract the votes and precinct name
        map_df(precincts, function(precinct) {
          precinct_name <- xml_attr(precinct, "name")
          votes <- xml_attr(precinct, "votes") %>% as.numeric()
          
          tibble(
            contest = contest_name,
            candidate = candidate_name,
            vote_type = vote_type_name,
            precinct = precinct_name,
            votes = votes
          )
        })
      })
    })
  }
  
  # Apply the function to all contests
  election_data <- contests %>% map_df(parse_contest_votes)
  return(election_data)
}
```

```{r}
# Load in the current precinct map and precinct Name Equivalency files
map <- st_make_valid(st_zm(st_read("../data/maps/precincts_2024.shp")))
name.equivalents <-readRDS("Precinct_name_id_equiv2024.rds")

# Load in previous maps 
map2018 <- st_make_valid(st_zm(st_read("../data/maps/precincts_2018.shp"))) %>% 
 st_transform(., crs = st_crs(map))
map2020 <- st_make_valid(st_zm(st_read("../data/maps/precincts_2020.shp"))) %>% 
 st_transform(., crs = st_crs(map))
map2022 <- st_make_valid(st_zm(st_read("../data/maps/precincts_2022.shp"))) %>% 
 st_transform(., crs = st_crs(map))

# Load the XML file and parse into a dataframe
election.data.2018 <- parse_contest_results(read_xml("old elections/detail2018.xml"))
election.data.2020 <- parse_contest_results(read_xml("old elections/detail2020.xml"))
election.data.2022 <- parse_contest_results(read_xml("old elections/detail2022.xml"))

# Pick which election to compare turnout to
turnout.data.2020 <-read_xml("old elections/detail2020.xml") %>%
    xml_find_all("//Precincts/Precinct")

```

Get census population data for interpolating between

```{r}
block.total <- get_decennial(geography = "block", 
                 state = "Ohio",
                 county = "Hamilton",
                 variables = "P1_001N", 
                 year = 2020,
                 sumfile = "dhc",
                 geometry = TRUE) %>% 
  select(total.pop = value) %>% 
  st_transform(crs = st_crs(map))
```
## Functions

A function to get turnout from the xml file
```{r}

turnout.df <- function(turnout.data){
  # Create a data frame
  turnout.df <- data.frame(
    name = xml_attr(turnout.data.2020, "name"),
    totalVoters = as.integer(xml_attr(turnout.data, "totalVoters")),
    ballotsCast = as.integer(xml_attr(turnout.data, "ballotsCast")),
    voterTurnout = as.numeric(xml_attr(turnout.data, "voterTurnout")),
    percentReporting = as.numeric(xml_attr(turnout.data, "percentReporting")),
    stringsAsFactors = FALSE
  )
  return(turnout.df)
}
```

A function to summarize the overall results of a head-to-head race.

```{r}
results.df <- function(my.race, my.dem, my.rep, election_data) {
  ## Take the total election results and sum up votes for Presidential candidates
  vote_sums <- election_data  %>%
    filter(contest == my.race) %>% 
    group_by(candidate, vote_type) %>%
    summarize(total = sum(votes, na.rm = TRUE),.groups="keep") %>% 
    pivot_wider(names_from = vote_type, values_from = total) %>% 
    mutate(Overall = rowSums(across(everything()))) %>% 
    ungroup()
  
  column_sums <- vote_sums %>%
    select(where(is.numeric)) %>%
    summarize(across(everything(), ~ sum(.x, na.rm = TRUE)))
  vote_sums <- vote_sums %>%
    add_row(candidate = "Total", !!!column_sums)
  
  ## Pull out the Democratic and Republican candidates and calculate their percentages
  dem.row = vote_sums %>% filter(candidate == my.dem)
  rep.row = vote_sums %>% filter(candidate == my.rep)
  tot.row = vote_sums %>% filter(candidate == 'Total')
  
  output = rbind(round(100*dem.row[2:4]/tot.row[2:4],1),round(100*rep.row[2:4]/tot.row[2:4],1))
  rownames(output) = c(my.dem,my.rep)
  
  return(output)
}
```

A function to get head-to-head election results by the precinct

```{r}
precinct.results.df <- function(my.race, my.dem, my.rep, election.data) {
  ## Get the precinct-by-precinct election results for Democrat and Republican in head-to-head race
  early <- election_data  %>%
    filter(contest == my.race,vote_type=='Early') %>% 
    group_by(precinct) %>%
    summarize(early = sum(votes, na.rm = TRUE),.groups="keep") 
  
  election <- election_data  %>%
    filter(contest == my.race,vote_type=='Election') %>% 
    group_by(precinct) %>%
    summarize(election = sum(votes, na.rm = TRUE),.groups="keep") 
  
  overall <- election_data  %>%
    filter(contest == my.race) %>% 
    group_by(precinct) %>%
    summarize(overall = sum(votes, na.rm = TRUE),.groups="keep")   
 
  dem.early <- election_data  %>%
    filter(contest == my.race,candidate==my.dem,vote_type=='Early') %>% 
    group_by(precinct) %>%
    summarize(dem.early = sum(votes, na.rm = TRUE),.groups="keep") 
  
  dem.election <- election_data  %>%
    filter(contest == my.race,candidate==my.dem,vote_type=='Election') %>% 
    group_by(precinct) %>%
    summarize(dem.election = sum(votes, na.rm = TRUE),.groups="keep") 
  
  dem.overall <- election_data  %>%
    filter(contest == my.race,candidate==my.dem) %>% 
    group_by(precinct) %>%
    summarize(dem.overall = sum(votes, na.rm = TRUE),.groups="keep") 
  
  rep.early <- election_data  %>%
    filter(contest == my.race,candidate==my.rep,vote_type=='Early') %>% 
    group_by(precinct) %>%
    summarize(rep.early = sum(votes, na.rm = TRUE),.groups="keep") 
  
  rep.election <- election_data  %>%
    filter(contest == my.race,candidate==my.rep,vote_type=='Election') %>% 
    group_by(precinct) %>%
    summarize(rep.election = sum(votes, na.rm = TRUE),.groups="keep") 
  
  rep.overall <- election_data  %>%
    filter(contest == my.race,candidate==my.rep) %>% 
    group_by(precinct) %>%
    summarize(rep.overall = sum(votes, na.rm = TRUE),.groups="keep")   
    
  output <-
    left_join(early,election,by='precinct') %>% 
    left_join(overall,by="precinct") %>% 
    left_join(dem.early,by="precinct") %>% 
    left_join(dem.election,by="precinct") %>% 
    left_join(dem.overall,by="precinct") %>% 
    left_join(rep.early,by="precinct") %>% 
    left_join(rep.election,by="precinct") %>% 
    left_join(rep.overall,by="precinct")
    
  return(output)
  }
```

# 2020 Turnout

## Get precinct-by-precinct turnout

```{r}
my.year = "2020"
turnout_data = turnout.data.2020
my.map = map2020

prec.turnout <- turnout.df(turnout_data)

mapANDresults <- left_join(my.map,prec.turnout,by=c("PRECINCT"="name"))

interpolated.results <- interpolate_pw(
  from = st_make_valid(mapANDresults),
  to = st_make_valid(map),
  to_id = "PRECINCT",
  extensive = TRUE,
  weights = st_make_valid(block.total),
  weight_column = "total.pop",
  crs = st_crs(map)
) 

m = min(interpolated.results$ballotsCast/interpolated.results$totalVoters)
M = max(interpolated.results$ballotsCast/interpolated.results$totalVoters)
interpolated.results %>% 
  ggplot(aes(fill=ballotsCast/totalVoters)) +
  geom_sf()+
  labs(title = paste('Voter Turnout',my.year,sep = ", "),
       subtitle = "",
       fill = "", 
       caption = "")+
  scale_fill_gradientn(colours=brewer.pal(n=10,name="Spectral"),na.value = "transparent",
                           breaks=c(0,.25,.50,.75,1),labels=c("0%","25%","50%","75%","100%"),
                           limits=c(m,M))+
  theme_void()

print(overall.results)



```


# 2020 Contests and Candidates

Print out all Contest/Candidates to be able to pull appropriate data

```{r}
print("-------2020-------")
print(unique(election.data.2020$contest))
print(unique(election.data.2020$candidate))
```

## Biden vs. Trump

```{r}
my.year = "2020"
my.race = "For President and Vice President"
my.dem = "Joseph R. Biden / Kamala D. Harris"
my.rep = "Donald J. Trump / Michael R. Pence"
election_data = election.data.2020
my.map = map2020

overall.results <- results.df(my.race,my.dem,my.rep,election_data)

prec.results <- precinct.results.df(my.race,my.dem,my.rep,election_data)
mapANDresults <- left_join(my.map,prec.results,by=c("PRECINCT"="precinct"))

interpolated.results <- interpolate_pw(
  from = st_make_valid(mapANDresults),
  to = st_make_valid(map),
  to_id = "PRECINCT",
  extensive = TRUE,
  weights = st_make_valid(block.total),
  weight_column = "total.pop",
  crs = st_crs(map)
) 

interpolated.results %>% 
  mutate(dem.prop = dem.overall/overall) %>%
  ggplot(aes(fill=dem.prop)) +
  geom_sf()+
  labs(title = paste(my.race,my.year,sep = ", "),
       subtitle = paste(my.dem, 'vs',my.rep,sep=' '),
       fill = "", 
       caption = "")+
  scale_fill_gradientn(colours=brewer.pal(n=10,name="RdBu"),na.value = "transparent",
                           breaks=c(0,.25,0.5,.75,1),labels=c("100% Rep","","50%/50%","","100% Dem"),
                           limits=c(0,1))+
  theme_void()

print(overall.results)
```
## Rucker vs. Deters

```{r}
my.year = "2020"
my.race = "For Prosecuting Attorney" 
my.dem = "Fanon A. Rucker"
my.rep = "Joseph T. Deters"
election_data = election.data.2020
my.map = map2020

overall.results <- results.df(my.race,my.dem,my.rep,election_data)

prec.results <- precinct.results.df(my.race,my.dem,my.rep,election_data)
mapANDresults <- left_join(my.map,prec.results,by=c("PRECINCT"="precinct"))

interpolated.results <- interpolate_pw(
  from = st_make_valid(mapANDresults),
  to = st_make_valid(map),
  to_id = "PRECINCT",
  extensive = TRUE,
  weights = st_make_valid(block.total),
  weight_column = "total.pop",
  crs = st_crs(map)
) 

interpolated.results %>% 
  mutate(dem.prop = dem.overall/overall) %>%
  ggplot(aes(fill=dem.prop)) +
  geom_sf()+
  labs(title = paste(my.race,my.year,sep = ", "),
       subtitle = paste(my.dem, 'vs',my.rep,sep=' '),
       fill = "", 
       caption = "")+
  scale_fill_gradientn(colours=brewer.pal(n=10,name="RdBu"),na.value = "transparent",
                           breaks=c(0,.25,0.5,.75,1),labels=c("100% Rep","","50%/50%","","100% Dem"),
                           limits=c(0,1))+
  theme_void()

print(overall.results)
```
# 2022 Contests and Candidates

```{r}
print("-------2022-------")
print(unique(election.data.2022$contest))
print(unique(election.data.2022$candidate))
```
## Ryan vs. Vance

```{r}
my.year = "2022"
my.race = "For U.S. Senator"
my.dem = "Tim Ryan"
my.rep = "JD Vance"
election_data = election.data.2022
my.map = map2022

overall.results <- results.df(my.race,my.dem,my.rep,election_data)

prec.results <- precinct.results.df(my.race,my.dem,my.rep,election_data)
mapANDresults <- left_join(my.map,prec.results,by=c("NAME"="precinct"))

interpolated.results <- interpolate_pw(
  from = st_make_valid(mapANDresults),
  to = st_make_valid(map),
  to_id = "PRECINCT",
  extensive = TRUE,
  weights = st_make_valid(block.total),
  weight_column = "total.pop",
  crs = st_crs(map)
) 

interpolated.results %>% 
  mutate(dem.prop = dem.overall/overall) %>%
  ggplot(aes(fill=dem.prop)) +
  geom_sf()+
  labs(title = paste(my.race,my.year,sep = ", "),
       subtitle = paste(my.dem, 'vs',my.rep,sep=' '),
       fill = "", 
       caption = "")+
  scale_fill_gradientn(colours=brewer.pal(n=10,name="RdBu"),na.value = "transparent",
                           breaks=c(0,.25,0.5,.75,1),labels=c("100% Rep","","50%/50%","","100% Dem"),
                           limits=c(0,1))+
  theme_void()

print(overall.results)
```
## Whaley vs. DeWine

```{r}
my.year = "2022"
my.race = "For Governor and Lieutenant Governor"
my.dem = "Nan Whaley and Cheryl L. Stephens"
my.rep = "Mike DeWine and Jon Husted"
election_data = election.data.2022
my.map = map2022

overall.results <- results.df(my.race,my.dem,my.rep,election_data)

prec.results <- precinct.results.df(my.race,my.dem,my.rep,election_data)
mapANDresults <- left_join(my.map,prec.results,by=c("NAME"="precinct"))

interpolated.results <- interpolate_pw(
  from = st_make_valid(mapANDresults),
  to = st_make_valid(map),
  to_id = "PRECINCT",
  extensive = TRUE,
  weights = st_make_valid(block.total),
  weight_column = "total.pop",
  crs = st_crs(map)
) 

interpolated.results %>% 
  mutate(dem.prop = dem.overall/overall) %>%
  ggplot(aes(fill=dem.prop)) +
  geom_sf()+
  labs(title = paste(my.race,my.year,sep = ", "),
       subtitle = paste(my.dem, 'vs',my.rep,sep=' '),
       fill = "", 
       caption = "")+
  scale_fill_gradientn(colours=brewer.pal(n=10,name="RdBu"),na.value = "transparent",
                           breaks=c(0,.25,0.5,.75,1),labels=c("100% Rep","","50%/50%","","100% Dem"),
                           limits=c(0,1))+
  theme_void()

print(overall.results)
```
# 2018 Contest and Candidates

```{r}
print("-------2018-------")
print(unique(election.data.2018$contest))
print(unique(election.data.2018$candidate))
```
## Brown vs. Renacci

```{r}
my.year = "2018"
my.race = "For U.S. Senator"
my.dem = "Sherrod Brown"
my.rep = "Jim Renacci"
election_data = election.data.2018
my.map = map2018

overall.results <- results.df(my.race,my.dem,my.rep,election_data)

prec.results <- precinct.results.df(my.race,my.dem,my.rep,election_data)
mapANDresults <- left_join(my.map,prec.results,by=c("PRECINCT"="precinct"))

interpolated.results <- interpolate_pw(
  from = st_make_valid(mapANDresults),
  to = st_make_valid(map),
  to_id = "PRECINCT",
  extensive = TRUE,
  weights = st_make_valid(block.total),
  weight_column = "total.pop",
  crs = st_crs(map)
) 

interpolated.results %>% 
  mutate(dem.prop = dem.overall/overall) %>%
  ggplot(aes(fill=dem.prop)) +
  geom_sf()+
  labs(title = paste(my.race,my.year,sep = ", "),
       subtitle = paste(my.dem, 'vs',my.rep,sep=' '),
       fill = "", 
       caption = "")+
  scale_fill_gradientn(colours=brewer.pal(n=10,name="RdBu"),na.value = "transparent",
                           breaks=c(0,.25,0.5,.75,1),labels=c("100% Rep","","50%/50%","","100% Dem"),
                           limits=c(0,1))+
  theme_void()

print(overall.results)
```


