---
title: "DSCI 210: Bells and Whistles"
author: "David Gerberry"
output: html
---

Picking up from our last module.


```{r setup, message=FALSE,warning=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(sf)
library(readxl)
library(RColorBrewer)
library(tidycensus)

```


# Bells and Whistles for maps

## Dot-density maps

Let's go back to a previous map of the results of the Trump/Biden race in 2020. This type of map is call a chlorepleth map (i.e. it's where you color in parts of map based on your data). These are wonderful maps, but no visualization is perfect.  One issue with chloropleth maps in general is that they *DO NOT* show us the way that population density varies.  Looking at this map, we might think Trump got more votes than Biden more of the map is red than blue. This is far from the case because a lot of the red parts of the map of much less densely populated.  Remember, land doesn't vote.  People vote.

```{r}

map2020 <- st_zm(st_read("../data/maps/precincts_2020.shp"))
results2020 <- read_excel("../data/election results/G20_Official_Canvass.xlsx", 
                       sheet = "Candidates",skip=1)
mapANDresults2020 <-
  left_join(map2020, results2020, by = c("PRECINCT" = "PRECINCT"))

mapANDresults2020 %>% 
  mutate(Biden.prop = `Biden & Harris  (Dem)`/( `Biden & Harris  (Dem)`+ `Trump & Pence       (Rep)`)) %>%
  ggplot(aes(fill=Biden.prop)) +
  geom_sf()+
  labs(title = "2020 Presidential Election", 
       subtitle = "Joe Biden vs Donald Trump",
       fill = "Vote for \nBiden (%)", 
       caption = "")+
  scale_fill_gradientn(colours=brewer.pal(n=10,name="RdBu"),na.value = "transparent",
                           breaks=c(0,.25,0.5,.75,1),labels=c("0%","25%","50%","75%","100%"),
                           limits=c(0,1))
```

One map that tries to address this is what's called a "dot-density" map.  Rather than coloring in an entire unit, the data is visualized as small dots and the denisty of the dots is determined by how many data points are in that area.

```{r}

trumpbiden <-
  mapANDresults2020 %>% 
  select(`Trump & Pence       (Rep)`,`Biden & Harris  (Dem)`) %>% 
  pivot_longer(cols =c("Trump & Pence       (Rep)","Biden & Harris  (Dem)"), names_to="candidate",values_to = "votes")

trumpbiden_dots <- as_dot_density(
  st_make_valid(trumpbiden),
  value = "votes",
  values_per_dot = 100,
  group = "candidate"
)

ggplot() +
  geom_sf(data = trumpbiden,
          fill = "white",
          color = "grey") +
  geom_sf(data = trumpbiden_dots,
          aes(color = candidate),
          size = 0.01) +
   scale_color_manual(values =  c("Trump & Pence       (Rep)" = "red", "Biden & Harris  (Dem)" = "blue"))+
  theme_void()

```


## mapview
Another weakness of our maps so far is that you really need to know Hamilton County (or Cincinnati) to understand what you're looking. For example, where is Xavier on the map above?
A way we can address this is the "mapview" package which will put the maps that we're making over top of Google Map.  It's OpenStreetMap, which is a free service, but you get the idea.

```{r}
library(mapview)

mapview(mapANDresults2020,zcol = "PERCENT",alpha=.05)

```

## Mapview and dot-density together

```{r}
mapview(trumpbiden_dots, zcol = "candidate", col.regions = c("blue", "red"), cex = 2, alpha = 0.7)


```

## Adding labels, text, to maps

So there are infinitely many cool things you can do with maps.  On the other hand, there is a special place for just a regular old image of map because most of our presentations are in PowerPoint (or something similar). So now let's focus on ways to add extra detail to these kinds of maps.

For example, say we want to pull out a particular municipality in the county, or let's say that we wanted to plot a particular point on our maps.

```{r}


mapANDresults2020 %>% 
  mutate(Biden.prop = `Biden & Harris  (Dem)`/( `Biden & Harris  (Dem)`+ `Trump & Pence       (Rep)`)) %>%
  ggplot(aes(fill=Biden.prop)) +
  geom_sf()+
  labs(title = "2020 Presidential Election", 
       subtitle = "Joe Biden vs Donald Trump",
       fill = "Vote for \nBiden (%)", 
       caption = "")+
  scale_fill_gradientn(colours=brewer.pal(n=10,name="RdBu"),na.value = "transparent",
                           breaks=c(0,.25,0.5,.75,1),labels=c("0%","25%","50%","75%","100%"),
                           limits=c(0,1))

### find which precincts in our dataframe are the ones from Norwood and the ones from Hyde Park with grepl() command
###  ?grepl    - to see its help page



grepl('NORW',mapANDresults2020$PRECINCT)
norwood = st_union(mapANDresults2020[grepl('NORW',mapANDresults2020$PRECINCT),])

cincy.neighborhoods <- st_zm(st_read("../data/maps/snabnd_2010.shp"))
hydepark  = cincy.neighborhoods[grepl('Hyde Park',cincy.neighborhoods$SNA_NAME),]




mapANDresults2020 %>% 
  mutate(Biden.prop = `Biden & Harris  (Dem)`/( `Biden & Harris  (Dem)`+ `Trump & Pence       (Rep)`)) %>%
  ggplot() +
  geom_sf(aes(fill=Biden.prop))+
  labs(title = "2020 Presidential Election", 
       subtitle = "Joe Biden vs Donald Trump",
       fill = "Vote for \nBiden (%)", 
       caption = "")+
  scale_fill_gradientn(colours=brewer.pal(n=10,name="RdBu"),na.value = "transparent",
                           breaks=c(0,.25,0.5,.75,1),labels=c("0%","25%","50%","75%","100%"),
                           limits=c(0,1)) +
  geom_sf(data = norwood,col='orange',fill=NA,lwd=1.5)+
  geom_sf_label(data=norwood,aes(label = "Norwood"),cex=2.5,position = position_nudge(x=0,y=0))+
  geom_sf(data = hydepark,col='pink',fill=NA,lwd=1.5)+
  geom_sf_label(data=hydepark,aes(label = "Hyde Park"),cex=2.5,position = position_nudge(x=.02,y=-.02))



```


