---
title: "DSCI 210: Lists of Registered Voters"
author: "David Gerberry"
output: html
---

Loading necessary packages and such.

```{r setup, message=FALSE,warning=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(sf)
library(readxl)
library(RColorBrewer)
library(tidycensus)

```


# Lists of registered voters

## Hamilton County Board of Elections

The Hamilton County Board of Elections (BOE) website is our main source of data for our campaign work. Under the "Campaign & Media Tools" tab, the BOE provides multiple useful data sets.  The `Registered Voter List' is updated daily in leading up to an election (see <https://votehamiltoncountyohio.gov/campaign-media/>). Voter registration closed one month before an election.  From that time on, the list of registered voters is complete and finalized.

```{r}
voters.boe <-read_csv("../data/registered voters/VoterListExport-20240917-pn.csv")
```


## Ohio Secretary of State

The Secretary of State (SOS) is the highest level administrator in charge of elections in Ohio.  From the SOS website, we can download registered voter lists as well (<https://www6.ohiosos.gov/ords/f?p=111:1>).  We can do this by county, the whole state, or a variety of electoral districts. For now, let's look at registered voters in Hamilton County.

```{r}
voters.sos <-read_csv("../data/registered voters/HAMILTON.csv")
```

## Why download both of these?

Let's take some notes on what's different between these two sets, (better/worse)
- more elections in Sec of State's data set
- fun
- different labels, different columns could be useful



# What can we do with the Registered Voter List?  ANS:  Data Science!

## What are the most democratic/republican areas

First, we can calculate the proportion of registered voters in each precinct that are Rep/Dem/Unaffiliated

```{r}
voter.summary <- 
voters.sos %>%
  mutate(PARTY_AFFILIATION = replace_na(PARTY_AFFILIATION, 'U')) %>% 
  group_by(PRECINCT_NAME) %>%
  summarize(total_voters = n(),
            un_voters = sum(PARTY_AFFILIATION == "U"), 
            rep_voters = sum(PARTY_AFFILIATION == "R"), 
            dem_voters = sum(PARTY_AFFILIATION == "D")) %>%
  mutate(dem.prop.of.reg = dem_voters / (dem_voters+rep_voters)) %>% 
  mutate(rep.prop.of.reg = rep_voters / (dem_voters+rep_voters)) %>% 
  mutate(unaffiliated.prop = un_voters / total_voters)
```


Now, let's see what are the most conservative/progressive areas of Hamilton County.

```{r}

poo = voters.sos %>%
  mutate(area = coalesce(TOWNSHIP,VILLAGE,WARD,CITY)) %>% 
  group_by(area) %>% 
  mutate(PARTY_AFFILIATION = replace_na(PARTY_AFFILIATION, 'U')) %>% 
  summarize(total_voters = n(),
            un_voters = sum(PARTY_AFFILIATION == "U"), 
            rep_voters = sum(PARTY_AFFILIATION == "R"), 
            dem_voters = sum(PARTY_AFFILIATION == "D")) %>%
  mutate(dem.prop.of.aff = dem_voters / (dem_voters+rep_voters)) %>% 
  mutate(rep.prop.of.aff = rep_voters / (dem_voters+rep_voters)) %>% 
  mutate(unaffiliated.prop = un_voters / total_voters)


most_dem <-poo %>% 
  arrange(desc(dem.prop.of.aff))
```

## It would be cool if we can map precincts by the proportion of unaffiliated voters


```{r}
map2024 <- st_zm(read_sf("../data/maps/precincts_2024.shp"))
mapANDvoters2024 <-
  left_join(map2024, voter.summary, by = c("PRC_NAME" = "PRECINCT_NAME"))

mapANDvoters2024 %>% 
  ggplot(aes(fill=unaffiliated.prop)) +
  geom_sf()+
  labs(title = "2024 Proportion of Unaffiliated Voters", 
       subtitle = "",
       fill = "", 
       caption = "")+
  scale_fill_gradientn(colours=brewer.pal(n=10,name="Spectral"),na.value = "transparent",
                           breaks=c(0,.25,0.5,.75,1),labels=c("0%","25%","50%","75%","100%"),
                           limits=c(0,1))


```




## Let's actually go into the List of Registered Voters and find people we may want to target with certain messages.


```{r}

election.date = "2024-11-05"

## start by only looking at the start of the data frame.  It will make our calculations run in a more reasonable amount of time.
## we'll do the first 5,000 rows only. 

# df = voters.sos 
df = voters.sos[1:5000,]

df <- df %>%  
   rename_with(~ gsub("-", ".", .))  

temp = df %>% 
  select(LAST_NAME,
         FIRST_NAME,
         DATE_OF_BIRTH,
         REGISTRATION_DATE,
         `PRIMARY.03/07/2000`,
         `GENERAL.11/07/2000`,
         `SPECIAL.05/08/2001`,
         `GENERAL.11/07/2000`)

temp <- temp %>%
  mutate(Age = as.numeric(difftime(as.Date(election.date), as.Date(DATE_OF_BIRTH), units = "weeks")) / 52.25)


check_after_registration <- function(election_col, election_date_col, registration_date) {
  # Check if election date is after registration date
  if_else(as.Date(election_date_col) < registration_date, election_col, 'o')
}

election_cols <- temp %>%
  select(starts_with("PRIMARY") | starts_with("GENERAL") | starts_with("SPECIAL")) %>%
  colnames()


for (X in election_cols){
  elect.date = str_extract(X,"(?<=\\.).*") %>%
  as.Date(format = "%m/%d/%Y")
  
  temp[[X]] = check_after_registration(temp[[X]],elect.date,temp$REGISTRATION_DATE)
}



# Extract the dates from those column names
election_dates <- election_cols %>%
  str_extract("(?<=\\.).*") %>%
  as.Date(format = "%m/%d/%Y")

# View the election columns with their extracted dates
tibble(election_cols, election_dates)



poo <- df %>%
  rowwise() %>%
  mutate(
    D_count = sum(across(47:127, ~ . == "D"),na.rm = T),
    R_count = sum(across(47:127, ~ . == "R"),na.rm = T)
  ) %>%
  ungroup()

```